---
title: "Data624_HW1"
author: "Alexis Mekueko"
date: "9/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r load-packages, results='hide',warning=FALSE, message=FALSE, echo=FALSE}

#library(tidyverse) #loading all library needed for this assignment
#remove.packages(tidyverse)
#library(openintro)
#library(lahman) #database for baseball
library(caret)
library(knitr)
#library(markdown)
#library(rmarkdown)
library(dplyr)
library(tidyr)
library(naniar)
library(reshape)
library(ggplot2)
library(qqplotr)
library(stats)
library(statsr)
library(GGally)
library(pdftools)
library(correlation)
library(Metrics)
library(e1071)
library(rocc)
library(pROC)
library(plm)
library(car)
#library(VIF)
#library(MASS)
#library(AICcmodavg)
#library(gridExtra)
#library(ggpubr)
#library(glmulti)

#install.packages("datarobot", dependencies = TRUE)

library(datarobot)
#install.packages("fpp3", dependencies = TRUE)
library(fpp3)
#install.packages("USgas", dependencies = TRUE)
library(USgas)

```



[Github Link](https://github.com/ATM1000k/Data608_HW1)
[Web Link](https://rpubs.com/amekueko/807900)




## Exercise 1

Use the help function to explore what the series gafa_stock, PBS, vic_elec and pelt represent.Use autoplot() to plot some of the series in these data sets. What is the time interval of each series?

```{r, echo=FALSE}

#View(inc)
#head(inc)
#summary(inc)
#help(gafa_stock)


#??gafa_stock
#help("gafa_stock")
print("Historical stock prices from 2014-2018 for Google, Amazon, Facebook and Apple. All prices are in $USD.")
#??PBS
#help(PBS)
print("PBS: maptools::SpatialLines2PolySet		Convert sp line and polygon objects to PBSmapping PolySet objects pbapply::pbapply		Adding Progress Bar to '*apply' Functions plm::pbsytest		Bera, Sosa-Escudero and Yoon Locally-Robust Lagrange Multiplier Tests for Panel Models and Joint Test by Baltagi and Liraster::readStart		Helper functions for programming splines::predict.bSpline		Evaluate a Spline at New Values of x")
#??pelt
#help(pelt)
print("Hudson Bay Company trading records for Snowshoe Hare and Canadian Lynx furs from 1845 to 1935. This data contains trade records for all areas of the company.")
#??vic_elec
#help("vic_elec")
print("Half-hourly electricity demand for Victoria, Australia")
#??mtcars


#ConnectToDataRobot(endpoint = "YOUR-ENDPOINT-HERE", token = "YOUR-API_TOKEN-HERE")
#GetProjectStatus(data(PBS))
autoplot(gafa_stock, Volume)
#max(gafa_stock$Date)
#min(gafa_stock$Date)
print("The timeline for gafa_stock series on volume stock goes from 2014-01-02 to 2018-12-31")

autoplot(PBS,Scripts)
#min(PBS$Month)

print("The PBS/Cost/Scripts didn't auotoplot but the timeline goes from July 1991 to June 2008")



autoplot(vic_elec, Demand)
#min(vic_elec$Time)
print("The autoplot is almost identical each year. The timeline for vic_elec series on power demand goes from 2012-01-01 AEDT to 2014-12-31 23:30:00 AEDT")
 
autoplot(pelt, Hare)
#min(pelt$Year)
print("The timeline for pel/Hare series goes from 1845 to 1935")


```

## Use filter() to find what days corresponded to the peak closing price for each of the four stocks in gafa_stock.

```{r }

view(gafa_stock)

sum(is.na(gafa_stock))

gafa_stock_close <- gafa_stock %>%
  dplyr::select(Symbol,Date,Close) %>%
  group_by(Symbol)%>%
  filter(Close == max(Close)) %>%
  arrange(desc(Close))

gafa_stock_close

  #%>%filter(complete.cases(.))


```

## Exercise 3

Download the file tute1.csv from the book website, open it in Excel (or some other spreadsheet application), and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

a-You can read the data into R with the following script:

```{r}

tute1 <- readr::read_csv("tute1.csv")
View(tute1)
sum(is.na(tute1))


```

b-Convert the data to time series

```{r }

mytimeseries <- tute1 %>%
  mutate(Quarter = yearmonth(Quarter)) %>%
  as_tsibble(index = Quarter)


```

c-Construct time series plots of each of the three series

```{r }

mytimeseries %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(x = Quarter, y = value, colour = name)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y")

```

Check what happens when you don’t include facet_grid().


```{r }

mytimeseries %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(x = Quarter, y = value, colour = name)) +
  geom_line() 

```
### The only difference is the groupinng. I think it would be redondant to use facet_grid if the the three variables have close range values.


## Exercise 4

The USgas package contains data on the demand for natural gas in the US.

a-Install the USgas package.
b-Create a tsibble from us_total with year as the index and state as the key.
c-Plot the annual natural gas consumption by state for the New England area (comprising the states of Maine, Vermont, New Hampshire, Massachusetts, Connecticut and Rhode Island).


```{r}


view(us_total)
glimpse(us_total)

us_total1 <- us_total %>%
  as_tibble(index = year, key = state)%>%
  filter(state == 'Connecticut' | state == 'Maine' | state == 'Massachusetts' | state == 'New Hampshire' | state == 'Rhode Island' | state == 'Vermont')#%>%

head(us_total1)

ggplot(data= us_total1, aes(x = year, y = y, col = state)) + 
  geom_line()+
   labs(x='Year', y="Natural Gas Consumption", title='Annual Natural Gas Consumption by State Region: case of New England Region')



```

## Exercise 5

a-Download tourism.xlsx from the book website and read it into R using readxl::read_excel().
b-Create a tsibble which is identical to the tourism tsibble from the tsibble package.

```{r}

tourism <- readxl::read_excel("tourism.xlsx")
View(tourism)
#colSums(is.na(tourism))%>% kable()
sum(is.na(tourism))
# tourism1 <- tourism %>%
#   mutate(Quarter = yearquarter(as.Date(tourism$Quarter)))%>%
#   as_tsibble( index = Quarter, key = c(Region, State, Purpose))

tourism$Quarter <- yearquarter(as.Date(tourism$Quarter))
glimpse(tourism)
tourism1 <- tourism %>%
   as_tsibble( index = Quarter, key = c(Region, State, Purpose))

```

c-Find what combination of Region and Purpose had the maximum number of overnight trips on average.


```{r}

tourism2 <- select( tourism, Region,Purpose, Trips) #something wrong with select....I think I was calling tourism1


# tourism2c <- tourism2 %>%
#   #dplyr::select(Region,Purpose, Trips) %>% #something wrong with select....
#   group_by(Region,Purpose)%>%
#   summarise(MeanTrip == mean(Trips)) %>%
#   filter(MeanTrip == max(MeanTrip))%>%
#   arrange(desc(MeanTrip))
# tourism2c
print("running into error running the above code, reason might be the datatype of Quarter, so I went back to adjust the datatype")

#rlang::last_error()
#rlang::last_trace()

```

d-Create a new tsibble which combines the Purposes and Regions, and just has total trips by State.


```{r}

tourism3 <- tourism %>%
  #dplyr::select(Region,Purpose, Trips) %>%
  group_by(Region, Purpose)
  tourism1$RegionPurpose <- paste(tourism1$Region, tourism1$Purpose)
  
# tourism3d <- tourism3 %>%
#   group_by(State) %>%
#   mutate(SumTripByState == sum(Trips)) %>%
#   as_tsibble( index = Quarter, key = c(Region, State, RPurpose))
#   #arrange(desc(MeanTrip))
# tourism3d
print("running into error running the above code, reason might be the datatype")


```

## Exercise 6

Monthly Australian retail data is provided in aus_retail. Select one of the time series as follows (but choose your own seed value):


Explore your chosen retail time series using the following functions:

Can you spot any seasonality, cyclicity and trend? What do you learn about the series?


```{r}

set.seed(53566)
myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))

autoplot(myseries)
gg_season(myseries)
gg_subseries(myseries)
gg_lag(myseries)

myseries %>%
  ACF(Turnover)%>%
  autoplot()
print("autoplot() gives a global view for the overall timeline, the trend line shows turnover never stop going wild. gg_season() allow a zoom in on the timeline to observe the turnover progression by month. gg_subseries() and gg_lag() goes a little more deeper than other function should one interested in looking up particular trend.")

```





